---
title: "Inferring Malaria Impact on Hemoglobin from Datasets, Including Those Where Women with Anaemia Are Excluded at Enrollment."
author: "Patrick Walker & Sequoia Leuba"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inferring Malaria Impact on Haemoglobin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Load necessary libraries and functions using `here`
library(here)
grav_cats <- 6
source(here("R", "load_libraries.R"))
source(here("R", "cubic_spline.R"))
source(here("R", "immune_functions.R"))
source(here("R", "prob_censor.R"))
source(here("R", "data_generation_functions.R"))
source(here("R", "define_params_functions.R"))
source(here("R", "log_likeli.R"))
source(here("R", "log_prior.R"))
source(here("R", "plot_functions.R"))

# Load pre-calculated infection history data
list_inf_histories <- readRDS(here::here("inst", "extdata", "site_inf_histories.rds"))
selected_inf_history <- list_inf_histories$site_inf_history[[1]]

# Generate synthetic data based on the selected infection history
synthetic_data_df <- generate_synthetic_data(site_inf_history = selected_inf_history)

```
This vignette applies the models developed in [Reference to Paper] to simulated data. We begin by validating our inferential framework for modelling the impact of malaria on haemoglobin levels during pregnancy, including the effect of acquired immunity over successive pregnancies. We then demonstrate the biases introduced by excluding women diagnosed with severe anaemia at enrolment and validate our approach for adjusting the framework to account for these biases, enabling us to correctly capture underlying trends.

## Inferring Patterns from Uncensored Data 

# Preparing Our Data 
Our study involved data on the HB concentrations at enrollment into a trial, infection status by PCR, gestational age at enrollment and gravidity. We also use ANC prevalence in primigravidae as a marker to determine our model-based estimates of patterns of prior exposure during pregnancy. The function generate_country_data_list_censor creates a list that involves both the data (here the synthetic data from data_generation.Rmd) and pre-calculated aggregated number of PCR samples and PCR positives in primigravidae within the dataset (to reduce computational time)

```{r generate synthetic data}
# Generate the uncensored data list
uncensored_data_list <- generate_country_data_list_censor(data_df = synthetic_data_df$data_df)

# Display the first few rows of the data
knitr::kable(head(uncensored_data_list$country_df))
## and the additional prevalence data added to the list
knitr::kable((uncensored_data_list$prev_data))
```

# defining parameters
Defining Parameters
Our approach uses the drjacoby package which requires specific templates for parameters. Here we use the function 'generate_df_params_non_censor' to set up a template of parameters required for this package. These parameters include:

+ Natural dynamics by gestational age (y_knots) and by gravidity (Gx_non_infect)
+ Dynamics of malaria impact by gestational age (mal_knots) and by immune status (shape/scale_hill)
+ Residual individual-level variability (HB_sigma)
+ A parameter for malaria prevalence at ANC in primigravidae to capture uncertainty in our sample.
```{r generate param_file}
# Generate parameter file for uncensored data
df_params_non_censor <- generate_df_params_non_censor()

# Display the parameter file
knitr::kable(df_params_non_censor)

```
Running the Model on Uncensored Data
With the data prepared and parameters defined, we now run our model on the uncensored dataset. This will help us validate our model's ability to capture the true underlying patterns without any biases introduced by data censorship.


```{r run_model}
run_noncensored_data <- run_mcmc(
  data = uncensored_data_list,
  df_params = df_params_non_censor,
  misc = list(
    gestage_min = 80,
    gestage_max = 200,
    grav_cats = grav_cats,
    cutoff = NULL,
    inf_history = selected_inf_history
  ),
  loglike = r_loglike,
  logprior = r_logprior_noncensor,
  burnin = 100,
  samples = 1000,
  pb_markdown = FALSE,
  chains = 1
)

```

```{r read model , echo = FALSE, eval = FALSE}
run_noncensored_data<-readRDS(here::here("inst", "extdata", "run_noncensored_data.rds"))
```
#Visualising the Results from Uncensored Data
After running the model, we can visualise the results to compare the model's predictions with the true synthetic data. This helps us ensure that the model correctly infers the impact of malaria on haemoglobin levels.
```{r generate plot noncensor model}
# Generate combined plot for non-censored data
combined_noncensored_plot <- generate_combined_plot(
  model = run_noncensored_data,
  fitted_data = uncensored_data_list$country_df,
  true_data = synthetic_data_df,
  site_inf_history = selected_inf_history
)


```

```{r  plot noncensor model,fig.width=10,fig.height=8}
# Display the combined plot
combined_noncensored_plot
```
##Introducing Bias: Inferring Patterns from Censored Data

In some of the studies used in the analysis, women with severe anaemia (HB < 7) are excluded from enrolment, introducing biases. Here, we simulate this scenario by censoring our synthetic data and then running the model without adjusting for this bias.
```{r  censor data}
# Censor women with HB < 7
cutoff <- 7
censored_data_list <- generate_country_data_list_noncensor(
  synthetic_data_df$data_df %>% filter(hb_level > cutoff)
)


```

```{r run unadjusted model}
# Run the model on censored data without adjustment
run_censored_data_incorrect <- run_mcmc(
  data = censored_data_list,
  df_params = df_params_non_censor,
  misc = list(
    gestage_min = 80,
    gestage_max = 200,
    cutoff = NULL,
    inf_history = selected_inf_history
  ),
  loglike = r_loglike,
  logprior = r_logprior_noncensor,
  burnin = 100,
  samples = 1000,
  pb_markdown = FALSE,
  chains = 1
)

```
#Visualising the Bias in the Results
We now visualise the results to see the impact of not adjusting for the bias introduced by excluding women with severe anaemia. The true impact of malaria upon haemoglobin is substantially under-estimated. 
```{r generate plot censored with unadjusted model}
# Generate combined plot for censored data without adjustment
combined_incorrect_censored_plot <- generate_combined_plot(
  model = run_censored_data_incorrect,
  site_inf_history = selected_inf_history,
  fitted_data = censored_data_list$country_df,
  true_data = synthetic_data_df
)
```
```{r  plot unadjusted model,fig.width=10,fig.height=8}
# Display the combined plot
combined_incorrect_censored_plot
```
##Correcting for Bias: Inferring Patterns from Censored Data with Adjustment

## preparing the inference
To account for the bias introduced by censoring, we adjust our model and re-run the analysis. Our approach assumes we know the number of individuals censored due to being anaemic at enrolment but not their gravidity distribution, nor their infection status. As a result, we have to make joint inference on the underlying true prevalence within each gravidity category, the true distribution of gravidity within our sample and the true underlying distribution of gestational age at enrollment. This requires we develop a new parameter file to reflect this.

```{r  generate define params censored  list}
# Generate parameter file for censored data
df_params_censor <- generate_df_params_censor()
knitr::kable(df_params_censor)
```

As our approach involves integrating over all of these factors to estimate the # of women who were anaemic at enrollment we need to do a great deal more pre-calculation to reduce computational intensity - in particular we need to calculate the observed distribution of gestational age 

```{r  generate censored data list}
# Generate data list for the censored data, including correction for bias
censored_data_list <- generate_country_data_list_censor(
  synthetic_data_df$data_df,
  cutoff = cutoff
)
knitr::kable(head(censored_data_list$country_df))

# This frequency of each unique day of gestational age, gravidity and malaria status within the dataset and an index variable to avoid having to loop over all possible combinations
knitr::kable(head(censored_data_list$counts_df))

# prevalence of malaria in primigravidae as before
knitr::kable(censored_data_list$prev_data)

# the distribution of counts by malaria status and gravidity
knitr::kable(censored_data_list$grav_mal_counts)

# the number censored due to being anaemic (Hb g/dL) of counts by malaria status and gravidity
knitr::kable(censored_data_list$censored_data)
```


```{r run adjusted model}
#We now run the model
# Run the model on censored data with bias correction
run_censored_data_correct <- run_mcmc(
  data = censored_data_list,
  df_params = df_params_censor,
  misc = list(
    gestage_min = 80,
    gestage_max = 200,
    cutoff = cutoff,
    inf_history = selected_inf_history
  ),
  loglike = r_loglike_w_censoring,
  logprior = r_logprior_censor,
  burnin = 200,
  samples = 1000,
  pb_markdown = FALSE,
  chains = 1
)
```
# Visualising the Corrected Results
Finally, we visualise the results from the corrected model to confirm that it accurately captures the true underlying patterns, even when women with severe anaemia are excluded.
```{r generate plot censored with correctly adjusted model}
# Generate combined plot for censored data with adjustment
combined_correct_censored_plot <- generate_combined_plot(
  model = run_censored_data_correct,
  site_inf_history = selected_inf_history,
  fitted_data = censored_data_list$country_df,
  true_data = synthetic_data_df
)
```

```{r  plot adjusted model,fig.width=10,fig.height=8}
# Display the combined plot
combined_correct_censored_plot
```

