---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Load necessary libraries and functions using `here`
library(here)
grav_cats<-6
source(here("R", "load_libraries.R"))
source(here("R", "cubic_spline.R"))
source(here("R", "immune_functions.R"))
source(here("R", "prob_censor.R"))
source(here("R", "data_generation_functions.R"))
source(here("R", "define_params_functions.R"))
source(here("R", "log_likeli.R"))
source(here("R", "log_prior.R"))
source(here("R", "plot_functions.R"))
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style


## Importing pre-calculated values
Our analysis uses an existing model to capture likely patterns of prior exposure by gravidity based upon patterns of fertility and Primigravid prevalence at first ANC

```{r read_inf_history, echo = TRUE}
list_inf_histories<-readRDS(here::here("inst", "extdata", "site_inf_histories.rds"))
knitr::kable(head(list_inf_histories$details))
```
Here we can see the closest available DHS survey year and admin1 unit the underlying fertility patterns are based upon. Selecting the first (representing the fertility patterns at the time of the trial of ISTp in Burkina Faso) we can look at it's structure. 

```{r show_inf_history_structure, echo = TRUE}
selected_inf_history<-list_inf_histories$site_inf_history[[1]]
names(selected_inf_history)
knitr::kable(head(selected_inf_history$sim_details))
knitr::kable(head(selected_inf_history$inf_mat_g2))
```

We can see that each fertility patterns contains a list of six dataframes summarising 1000 simulations of the model. 'sim_details' shows the EIRs used for each simulations the model ranging from $10^-0.5$ and $10^4$ sampled uniformally on the log-scale and primi_ANC_prev the resultant prevalence at first ANC in primgravid women. inf_mat_gX is then the distribution of levels of previous exposure in a previous pregnancy.

So if we want to see what the distribution of exposures of women infected during their second or third pregnancy in a setting with 10% prevalence by PCR at first ANC we use 'sim_details' as our look-up table and use that to access the correct row of 'inf_mat_g2' and 'inf_mat_g3'respectively

```{r show how to access correct element, echo = TRUE}
primi_prev<-0.1
 index <- which.min(abs(selected_inf_history$sim_details$primi_ANC_prev - primi_prev))
  selected_inf_history$inf_mat_g2[index,]
  selected_inf_history$inf_mat_g3[index,]
```

So for infected secundigravidae we'd expect around 25% to have been exposed to malaria in their first pregnancy. 
At first glance this may seem unintuitively high given prevalence was 10% at first ANC in their previous pregnancy. In part this is because the model capture cumulative prevalence from 12 weeks gestation (when the placenta is considered sufficiently viable to lead to substantial sequestration) to delivery but is largely due to heterongeneity in exposure within the community - women who are infected this pregnancy are more likely than average to be infected this one. 

To model how prior exposure mitigates the malaria-associated impact on hemoglobin, we use the Hill function, which is commonly used in biological contexts to describe saturation effects.
```{r hill_func, echo = TRUE,fig.width=6}

# Example usage of the Hill function
shape <- 2
scale <- 1
immunity_vector <- hill_func(0:20, shape, scale)



# Plot the Hill function
library(ggplot2)
ggplot(data.frame(i = 0:20, immunity = immunity_vector), aes(x = i, y = immunity)) +
  geom_line() +
  labs(title = "Immunity Vector Modeled by Hill Function",
       x = "Number of Prior Exposures",
       y = "Immunity Effect") +
  theme_minimal()
```

We can integrate the immunity vector over the distribution of exposure to get a weighted effect of immunity by gravidity for a given ANC primigravid prevalence using the function get_weight_impact_df.
```{r show how to generate immunity gravidity, echo = TRUE}
# Assuming primi_prev is given and selected_inf_history is loaded
imm_vect <- hill_func(0:20, shape = 2, scale = 3)
 get_weight_impact_df(
  inf_history = selected_inf_history,
  primi_prev = 0.1,
  imm_vect = imm_vect
)

  get_weight_impact_df(
  inf_history = selected_inf_history,
  primi_prev = 0.5,
  imm_vect = imm_vect
)

```
we can integrate over the distribution of prior exposure to get a measure of how 


synthetic_data<-generate_synthetic_data(site_inf_history=selected_inf_history)



Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
